Length Gauge Theory of Surface Second Harmonic Generation

### 2014-11-19 paper-v-0.tex
added new references and updated bib file

In recent years, surface nonlinear optical spectroscopies and in 
particular, surface second-harmonic generation SSHG have evolved as useful 
nondestructive and noninvasive tool to study properties of surfaces 
and interfaces: atomic structure, phase transitions, and adsorption of 
atoms,1–11{\color{\chonn} [incluye estas de mejiaPRB02 y combinalas con
  las tuyas y tambien pon la de salazarPRB14 y busca referencia para
  ??, checando q sean relevantes con SSHG]} 
among others. \cite{hughesPRB96, 
     guyot-sionnestPRB88, 
     lupkeSSR99, 
     downerPSSA01, 
     shenAPB99, 
     shenNAT89, 
     chenPRL81, 
    mendozaPRL98,downerSIA01,cazzanelliNM14}
Nowadays, SSHG spectroscopy is a crucial tool for research and development
in microelectronics,\cite{??} semiconductors,\cite{??} nanomaterials,\cite{??} and many more recent 
areas of great scientific and commercial interest.\cite{??}
The high surface sensitivity of SSHG spectroscopy is due to the
fact that, within the dipole approximation, the bulk SHG signal of
centrosymmetric materials is identically zero, and thus only the
surface, where the inversion symmetry is broken, can radiate. 
Much of the foundation of surface science has been built from
experiments involving emission or scattering of electrons from
surfaces. These require ultrahigh vacuum (UHV) environments and
provide no access to buried interfaces. 
However, SSHG
is compatibile with non-UHV ambients and has access to
interfaces buried beneath transparent overlayers. Even when applied to
surfaces in UHV, the light source and detectors can be aligned and
used outside the vacuum chamber.

@@

As it is well known
Density Functional Theory (DFT) within the Local Density Approximation
(LDA),\cite{??} the most 
used framework for \textit{ab initio} calculations,
underestimates the energy band gap of semiconductors. It is also well
understood that one has to include the many-body interaction to
correct for the underestimation of the gap. In this context the so
called GW approximation\cite{onidaRMP01} is known to correct the band
gap of most semiconductors\cite{cual}. However this could be a numerically very
expensive calculation, and thus one uses the much simpler scissors
operator scheme.\cite{?} This allows to ``open'' the DFT-LDA gap to
its correct experimental or GW value, when experiments are not available. 
In this context,
to correct for the underestimation of the energy band gap of semiconductors, 
Nastos et al.\cite{nastosPRB05} 
using the 
``length gauge'' or 
``$\mathbf{r}\cdot\mathbf{E}$ gauge'' 
showed how to correctly include in the SH susceptibility the many-body corrections through the
scissors operator.
Later,
Cabellos \textit{et al}.\cite{cabellosPRB09} 
elaborated a derivation of the ``velocity gauge'' or
``$\bfA\cdot\bfv$'' gauge properly including the scissors
operator, and proved  gauge 
invariance with respect to the length gauge.
From these works, it is clear the length gauge is a much better staring
point to obtain the surface SH (SSH) susceptibility, as it will be
elaborated in this article. Above only being valid for bulk semiconductors.

@@

a Kramers-Kronig transformation\cite{tacongePRB14} {\color{\chon} and}

@@

\subsection{Full-slab results}\label{results}

>>>>

In recent years, surface nonlinear optical spectroscopies and in 
particular, surface second-harmonic generation SSHG have evolved as useful 
nondestructive and noninvasive tool to study properties of surfaces 
and interfaces: atomic structure, phase transitions, and adsorption of 
atoms, among others.\cite{
    daumPRL93, 
    mcgilpOE94, 
    meyerPRL95, 
    powerPRL95, 
    godefroyAPL96, 
    hoferAPA96, 
    dadapPRB97, 
    bloembergenAPB99, 
    mcgilpSRL99, 
    suzukiAPB99, 
    mitchellSS01,
    hughesPRB96, 
    guyot-sionnestPRB88, 
    downerPSSA01, 
    shenAPB99, 
    shenNAT89, 
    chenPRL81, 
    mendozaPRL98,
    downerSIA01,
    cazzanelliNM14}
Nowadays, SSHG spectroscopy is a crucial tool for research and development
in microelectronics,\cite{zheltikovLP00} semiconductors,
\cite{lupkeSSR99}, nanomaterials,\cite{salazar-aparicioPRB14} 
and many more recent 
areas of great scientific and commercial interest.\cite{cazzanelliNM14}
The high surface sensitivity of SSHG spectroscopy is due to the
fact that, within the dipole approximation, the bulk SHG signal of
centrosymmetric materials is identically zero, and thus only the
surface, where the inversion symmetry is broken, can radiate. 
Much of the foundation of surface science has been built from
experiments involving emission or scattering of electrons from
surfaces. These require ultrahigh vacuum (UHV) environments and
provide no access to buried interfaces. 
However, SSHG
is compatibile with non-UHV ambients and has access to
interfaces buried beneath transparent overlayers. Even when applied to
surfaces in UHV, the light source and detectors can be aligned and
used outside the vacuum chamber.

@@

As it is well known
Density Functional Theory (DFT) within the Local Density Approximation
(LDA),\cite{kohnPR65} the most 
used framework for \textit{ab initio} calculations,
underestimates the energy band gap of semiconductors. It is also well
understood that one has to include the many-body interaction to
correct for the underestimation of the gap. In this context the so
called GW approximation\cite{onidaRMP02} is known to correct the band
gap of most semiconductors\cite{luceroJPCM12}. However this could be a numerically very
expensive calculation, and thus one uses the much simpler scissors
operator scheme.%%\cite{?} find this
This allows to ``open'' the DFT-LDA gap to
its correct experimental or GW value, when experiments are not available. 
In this context,
to correct for the underestimation of the energy band gap of semiconductors, 
Nastos et al.\cite{nastosPRB05} 
using the 
``length gauge'' or 
``$\mathbf{r}\cdot\mathbf{E}$ gauge'' 
showed how to correctly include in the SH susceptibility the many-body corrections through the
scissors operator.
Later,
Cabellos \textit{et al}.\cite{cabellosPRB09} 
elaborated a derivation of the ``velocity gauge'' or
``$\bfA\cdot\bfv$'' gauge properly including the scissors
operator, and proved  gauge 
invariance with respect to the length gauge.
From these works, it is clear the length gauge is a much better staring
point to obtain the surface SH (SSH) susceptibility, as it will be
elaborated in this article. Above only being valid for bulk semiconductors.

@@

a Kramers-Kronig transformation\cite{nicolasPRB14} {\color{\chon} and}

@@

\subsection{Full-slab results}\label{fsresults}
